Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Shanghai version \YellowPaperVersionNumber}}

### 2024-05-25 Paper.tex
eip-3651-alexprut-review
eip-3860-alexprut-review

\subsection{Execution}
\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{N({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
\nonumber {} & + \sum_{j = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \big( G_{\mathrm{accesslistaddress}} + \lVert T_{\mathbf{A}}[j]_{\mathbf{s}} \rVert G_{\mathrm{accessliststorage}} \big)
\end{align}
where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
$\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
$G$ is fully defined in Appendix \ref{app:fees}.

\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $N$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:

\begin{equation}
  N(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
\end{equation}

@@

Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{equation}
(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
\end{cases}
\end{equation}
where
\begin{align}
A^* & \equiv A^0 \quad \text{except} \\
A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\} \\
A^*_{\mathbf{a}} & \equiv \begin{cases}
a \cup T_{\mathrm{t}} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
a & \text{otherwise}
\end{cases} \\
a & \equiv A^0_{\mathbf{a}} \cup \{S(T)\} \cup \{H_c\}\cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
\end{align}
and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_{\mathrm{g}} - g_0
\end{equation}

@@

C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
G_{\mathrm{create}} + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{N(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

>>>>

\subsection{Execution}
\hypertarget{intrinsic_gas_g_0}{}We define intrinsic gas $g_0$, the amount of gas this transaction requires to be paid prior to execution, as follows:
\begin{align}
g_0 \equiv {} & \sum_{i \in T_{\mathbf{i}}, T_{\mathbf{d}}} \begin{cases} \hyperlink{G__txdatazero}{G_{\mathrm{txdatazero}}} & \text{if} \quad i = 0 \\ \hyperlink{G__txdatanonzero}{G_{\mathrm{txdatanonzero}}} & \text{otherwise} \end{cases} \\
\nonumber {} & + \begin{cases} \hyperlink{G__txcreate}{G_{\mathrm{txcreate}}} + \hyperlink{initcode_cost}{R({\lVert T_{\mathbf{i}} \rVert})} & \text{if} \quad T_{\mathrm{t}} = \varnothing \\ 0 & \text{otherwise} \end{cases} \\
\nonumber {} & + \hyperlink{G__transaction}{G_{\mathrm{transaction}}} \\
\nonumber {} & + \sum_{j = 0}^{\lVert T_{\mathbf{A}} \rVert - 1} \big( G_{\mathrm{accesslistaddress}} + \lVert T_{\mathbf{A}}[j]_{\mathbf{s}} \rVert G_{\mathrm{accessliststorage}} \big)
\end{align}
where $T_{\mathbf{i}},T_{\mathbf{d}}$ means the series of bytes of the transaction's associated data and initialisation EVM-code, depending on whether the transaction is for contract-creation or message-call.
$G_{\mathrm{txcreate}}$ is added if the transaction is contract-creating, but not if a result of EVM-code.
$\hyperlink{G__accesslistaddress}{G_{\mathrm{accesslistaddress}}}$ and $\hyperlink{G__accessliststorage}{G_{\mathrm{accessliststorage}}}$ are the costs of warming up account and storage access, respectively.
$G$ is fully defined in Appendix \ref{app:fees}.

\hypertarget{initcode_cost}{}We define the \textit{initcode cost function}, formally $R$, as the amount of gas that needs to be paid for each word of the initcode prior to executing the creation code of a new contract:

\begin{equation}
  R(x) \equiv \hyperlink{G__initcodeword}{G_{\mathrm{initcodeword}}} \times \lceil \frac{x}{32} \rceil
\end{equation}

@@

Evaluating $\boldsymbol{\sigma}_{\mathrm{P}}$ from $\boldsymbol{\sigma}_0$ depends on the transaction type; either contract creation or message call; we define the tuple of post-execution provisional state $\boldsymbol{\sigma}_{\mathrm{P}}$, remaining gas $g'$, accrued substate $A$ and status code $z$:
\begin{equation}
(\boldsymbol{\sigma}_{\mathrm{P}}, g', A, z) \equiv \begin{cases}
\hyperlink{lambda}{\Lambda}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), g, &\\ \quad\; p, T_{\mathrm{v}}, T_{\mathbf{i}}, 0, \varnothing, \top) & \text{if} \quad T_{\mathrm{t}} = \varnothing \\
\hyperlink{theta}{\Theta}_{4}(\boldsymbol{\sigma}_0, A^*, S(T), S(T), T_{\mathrm{t}}, &\\ \quad\; T_{\mathrm{t}}, g, p, T_{\mathrm{v}}, T_{\mathrm{v}}, T_{\mathbf{d}}, 0, \top) & \text{otherwise}
\end{cases}
\end{equation}
where
\begin{align}
A^* & \equiv A^0 \quad \text{except} \\
A^*_{\mathbf{K}} & \equiv \bigcup_{E \in T_{\mathbf{A}}} \big\{ \forall i < \lVert E_{\mathbf{s}} \rVert, i \in \mathbb{N}: \; (E_{\mathrm{a}}, E_{\mathbf{s}}[i]) \big\} \\
A^*_{\mathbf{a}} & \equiv \begin{cases}
a \cup T_{\mathrm{t}} & \text{if} \quad T_{\mathrm{t}} \neq \varnothing \\
a & \text{otherwise}
\end{cases} \\
a & \equiv A^0_{\mathbf{a}} \cup S(T) \cup H_c \cup_{E \in T_{\mathbf{A}}} \{ \hyperlink{access_list_entry}{E}_{\mathrm{a}} \}
\end{align}
and $g$ is the amount of gas remaining after deducting the basic amount required to pay for the existence of the transaction:
\begin{equation}
g \equiv T_{\mathrm{g}} - g_0
\end{equation}

@@

C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) + G_{\mathrm{copy}}\times\lceil\boldsymbol{\mu}_{\mathbf{s}}[3] \div 32\rceil & \text{if} \quad w = \text{\small EXTCODECOPY} \\
C_{\mathrm{aaccess}}(\boldsymbol{\mu}_{\mathbf{s}}[0] \bmod 2^{160}, A) & \text{if} \quad w \in W_{\mathrm{extaccount}}\\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1] & \text{if} \quad w = \text{\small LOG0} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG1} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+2G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG2} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+3G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG3} \\
G_{\mathrm{log}}+G_{\mathrm{logdata}}\times\boldsymbol{\mu}_{\mathbf{s}}[1]+4G_{\mathrm{logtopic}} & \text{if} \quad w = \text{\small LOG4} \\
C_\text{\tiny CALL}(\boldsymbol{\sigma}, \boldsymbol{\mu}, A) & \text{if} \quad w \in W_{\mathrm{call}} \\
C_\text{\tiny SELFDESTRUCT}(\boldsymbol{\sigma}, \boldsymbol{\mu}) & \text{if} \quad w = \text{\small SELFDESTRUCT} \\
G_{\mathrm{create}} + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small CREATE}\\
G_{\mathrm{create}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[2] \div 32 \rceil + \hyperlink{initcode_cost}{R(\boldsymbol{\mu}_{\mathbf{s}}[2])} & \text{if} \quad w = \text{\small \hyperlink{create2}{CREATE2}}\\
G_{\mathrm{keccak256}}+G_{\mathrm{keccak256word}} \times \lceil \boldsymbol{\mu}_{\mathbf{s}}[1] \div 32 \rceil & \text{if} \quad w = \text{\small KECCAK256}\\
G_{\mathrm{jumpdest}} & \text{if} \quad w = \text{\small JUMPDEST}\\
C_\text{\tiny SLOAD}(\boldsymbol{\mu}, A, I) & \text{if} \quad w = \text{\small SLOAD}\\
G_{\mathrm{zero}} & \text{if} \quad w \in W_{\mathrm{zero}}\\
G_{\mathrm{base}} & \text{if} \quad w \in W_{\mathrm{base}}\\
G_{\mathrm{verylow}} & \text{if} \quad w \in W_{\mathrm{verylow}}\\
G_{\mathrm{low}} & \text{if} \quad w \in W_{\mathrm{low}}\\
G_{\mathrm{mid}} & \text{if} \quad w \in W_{\mathrm{mid}}\\
G_{\mathrm{high}} & \text{if} \quad w \in W_{\mathrm{high}}\\
G_{\mathrm{blockhash}} & \text{if} \quad w = \text{\small \hyperlink{blockhash}{BLOCKHASH}}\\
\end{cases}
\end{equation}
\begin{equation}
w \equiv \begin{cases} I_{\mathbf{b}}[\boldsymbol{\mu}_{\mathrm{pc}}] & \text{if} \quad \boldsymbol{\mu}_{\mathrm{pc}} < \lVert I_{\mathbf{b}} \rVert\\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}
